---
title: "PML Project"
author: "Yagna"
date: "November 22, 2015"
output: html_document
---

##Loading data and required packages
```{r loading data and packages}
library(ggplot2)
library(caret)
library(randomForest)
library(gbm)
setwd('C:/Users/orugyag/Contacts/Desktop/WORK/Training/JHU_Data_Science_Specialization/Practical Machine Learning')
tr<-read.csv("pml-training.csv")
te<-read.csv("pml-testing.csv")
```

##Filtering the training and testing datasets
- Removing columns which aren't useful in the model fit (timestamp columns and window columns)
- Removing columns which have statistics of the data and retaining only columns with raw data (i.e. all columns with   avg/kurtosis/stddev/max/min etc.)

```{r filtering training dataset}
tr<-data.frame(tr)
cn<-colnames(tr)
statColInd<-c(grep("avg",cn),grep("kurtosis",cn),grep("skewness",cn),grep("max",cn),grep("min",cn),grep("amplitude",cn),grep("var",cn),grep("stddev",cn))
tr_sub<-tr[,-statColInd]
tr_sub2<-tr_sub[,-c(1,3,4,5,6,7)]
```

```{r filtering test dataset}
te<-data.frame(te)
cn_te<-colnames(te)
statColInd_te<-c(grep("avg",cn_te),grep("kurtosis",cn_te),grep("skewness",cn_te),grep("max",cn_te),grep("min",cn_te),grep("amplitude",cn_te),grep("var",cn_te),grep("stddev",cn_te))
te_sub<-te[,-statColInd_te]
te_sub2<-te_sub[,-c(1,3,4,5,6,7,60)]
```

##Boosting Model
- I fit a boosting algorithm with 10-fold cross validation to predict the classe

```{r boosting model}
set.seed(123)
boostFit <- train(classe ~ ., method = "gbm", data = tr_sub2, verbose = F, trControl = trainControl(method = "cv", number = 10))
plot(boostFit, ylim = c(0.8, 1))
```

The boosting algorithm generated a good model with accuracy = 0.9719
The expected out of sample error is 0.04

##Final Model and Prediction

```{r prediction}
tr_predict<-predict(boostFit,newdata = tr_sub2)
confusionMatrix(tr_predict,tr_sub2$classe)

te_predict<-as.character(predict(boostFit,newdata = te_sub2))
```

##Write prediction files

```{r write prediction files}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./prediction/problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(te_predict)
```